A Fundamental Measure Theory Functional for Hard-Sphere Contact Densities

### 2012-05-22 papers/contact/paper.tex

@@

The procedure for finding the symmetrically averaged correlation
function $g_\sigma^S(\rr)$ is less straight-forward than 
\begin{align}
  n_3^S(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-R(\rr))d\rr'
\end{align}
The hard-sphere radius $R$ in this weighted density has an unusual
position dependence:  it depends on the position at which we are
evaluating the weighted density, rather than the position of the hard
spheres themselves.  This has the effect of expanding spheres as they
touch the point $\rr$, a concept illustrated in
Figure~\ref{fig:contact}.  This enables a similar derivation to that
which gave us $g_\sigma^A(\rr)$:
\begin{figure}
\includegraphics[width=5cm]{figs/contact}
\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
  consist of those which just touch the point $\mathbf{r}$.}
\label{fig:contact}
\end{figure}
\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \\
  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
  &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
\end{align}
where the functional derivative $\frac{\delta A_{HS}}{\delta
  R(\mathbf{r})}$ in this case is computed with the symmetric
formulation of the fundamental measures.  Complete expressions for
$g_\sigma^S$ are included in Appendix~\ref{appendix:g-S}.\\

>>>>

In addition to the different densities in the denominators of Equations 
\ref{eq:g-A-exact-a} and \ref{eq:g-S-exact-a} there is also a difference 
between the treatment of the A and S cases in the derivatives 
$\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
and \ref{eq:g-S-exact}. The White Bear approximations for $A_{HS}$ 
that we use involve a number of weighted densities and for the derivatives 
of these we use two different conceptual approaches.  In the assymetric case, 
a change in $R(\mathbf{r})$ will refer to a change in the radius of a sphere 
that is located at the position $\mathbf{r}$.  The weighted densities that are 
affected by such a change will be those that are located a distance $R$ away 
from such a sphere.  We compute the assymetric derivatives In Appendix~\ref{appendix:g-A}.

@@

The procedure for finding the symmetrically averaged correlation
function $g_\sigma^S(\rr)$ is less straight-forward than 
\begin{align}
  n_3^S(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-R(\rr))d\rr'
\end{align}
The hard-sphere radius $R$ in this weighted density has an unusual
position dependence:  it depends on the position at which we are
evaluating the weighted density, rather than the position of the hard
spheres themselves.  This has the effect of expanding spheres as they
touch the point $\rr$, a concept illustrated in
Figure~\ref{fig:contact}.  This enables a similar derivation to that
which gave us $g_\sigma^A(\rr)$:
%% =======
%% For the computation of the symmetrically averaged correlation function 
%% we must formulate derivatives with respect to $R$ in a symmetric fashion.  
%% Conceptually we expand the radii of spheres that just touch a specific 
%% point of contact and only expand them outward at that point of contact.  
%% Therefore $R(\mathbf{r})$ in our calculations refers to the radii of 
%% spheres that touch $\mathbf{r}$ and $\delta R(\mathbf{r})$ refers to a 
%% change of these radii at only the point $\mathbf{r}$.  One can imagine 
%% a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell 
%% area protruding outward at $\mathbf{r}$.  Figure~\ref{fig:contact} 
%% illustrates such a change in $R$.  The full derivations are found in Appendix~\ref{appendix:g-S}

%% \begin{figure}
%% \includegraphics[width=5cm]{figs/n0}
%% \caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
%%   consist of those which just touch the point $\mathbf{r}$.}
%% \label{fig:n0}
%% \end{figure}

%% These qualitative concepts lead us to the derivative terms in $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$, 
%% which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the 
%% derivation of the correlatation function results when we recognize that 
%% instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$ 
%% as in the assymmetric case we will instead divide by $n_0(\mathbf{r})^2$ 
%% since we have symmetry around the point $\mathbf{r}$.

%% \begin{align}
%%   p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
%%     A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
%%   \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
%%     A_{HS}}{\delta R(\mathbf{r})} \\
%%   g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \label{eq:g-S-exact-a}\\
%%   g_\sigma^A(\rr) &= \frac{1}{n_0(\rr) n_0(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
%%     A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-S-exact}
%% \end{align}Complete expressions for $g_\sigma^S$ are included in
%% Appendix~\ref{appendix:g-S}.\\

%% \section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
%% We compare with the averaged correlation function of
%% Gross\cite{gross2009density}, which is of the asymmetrical variety
%% ($g_\sigma^A$), where the correlation function is averaged at location
%% of one of the spheres in concact.
%% \begin{align}
%%   g_\sigma^\text{Gross,A} &= \frac{1 - \frac12 n_{3'}}{\left(1 -
%%     n_{3'}\right)^3} \\
%%   n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
%% \end{align}
%% where $n_{3'}$ is an averaged
%% filling fraction defined in a manner analogous to the $n_3$ of FMT,
%% but with the hard-sphere radius replaced by the hard-sphere diameter,
%% so the integrals range over spheres that are twice as large.

%% >>>>>>> a989df3e529c6d7afbecef041c2a6afb3cb30c08
\begin{figure}
\includegraphics[width=5cm]{figs/contact}
\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
  consist of those which just touch the point $\mathbf{r}$.}
\label{fig:contact}
\end{figure}
\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \\
  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
  &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
\end{align}
where the functional derivative $\frac{\delta A_{HS}}{\delta
  R(\mathbf{r})}$ in this case is computed with the symmetric
formulation of the fundamental measures.  Complete expressions for
$g_\sigma^S$ are included in Appendix~\ref{appendix:g-S}.\\
